ArKanjo: a tool for detecting function-level Code Duplication in the Linux Kernel

### 2025-06-17 paper/paper.tex
Adding the Acknowledgment section
Reviewing the ArKanjo Tool section

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Luan Ícaro Pinto Arcanjo}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil 05508-220\\
Email: luanicaro@usp.br}
\and
\IEEEauthorblockN{David Tadokoro}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil 05508-220\\
Email: davidbtadokoro@usp.br}
\and
\IEEEauthorblockN{Paulo Meirelles}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil 05508-220\\
Email: paulormm@ime.usp.br}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Linux kernel’s massive scale (+28 M LoC, +20 K contributors) presents unique 
maintenance challenges. Surprisingly, code duplication remains a persistent issue 
in the kernel’s codebase, which can hinder its evolution and patching. Academic 
approaches often focus on pairwise comparison of code artifacts, which are not directly applied 
to comprehensive codebase analyses. Other existing free software tools explored in 
practice frequently suffer from limited functionality, such as primitive textual 
matching, prove too narrow in scope, or fail to deliver effective results on complex, 
large-scale codebases. Existing solutions generally fail to address the Linux kernel’s 
specific needs: (1) scalability to handle its size, (2) actionable results for 
developers, and (3) integration with kernel development workflows.
%
This paper presents \textit{ArKanjo}, a novel command-line tool for Linux kernel maintenance 
designed to detect and analyze function-level duplications. Released under the MIT 
license, ArKanjo employs a two-stage architecture consisting of a Preprocessor and 
a Query Responder that separates computationally intensive analysis from efficient 
querying for duplications within large codebases. Pivotal advantages of ArKanjo over 
existing solutions include: (1) optimization for C codebases with kernel-specific 
patterns; (2) preprocessing that enables rapid queries without redundant analysis; 
and (3) prioritization of duplicates that impact maintainability, such as copied 
buggy logic. 
%
We evaluate ArKanjo against real-world duplication cases in recent 
kernel versions, demonstrating its effectiveness in identifying problematic clones 
that generic tools often overlook. By identifying well-defined, manageable 
duplication instances, ArKanjo effectively lowers the barrier for new contributors, 
a capability evidenced by its role in guiding students to make their first code 
improvements to the kernel. ArKanjo offers immediate value to kernel maintainers 
and serves as a replicable model for clone detection in other large-scale free 
software projects.

@@

The Linux kernel is a foundational Free Software project 
critical to a significant portion of the world digital infrastructure. Maintaining 
the kernel is an enormous undertaking involving over 28 million lines of code 
and contributions from over twenty thousand developers. Within this context, code 
duplication persists as a significant challenge, a known harmful practice that 
negatively affects code readability and increases the likelihood of introducing
bugs~\cite{harmone,harmtwo}. This issue is particularly acute in the 
kernel device drivers, which comprise over 66\% of the source code~\cite{marcelo}. 
For instance, maintainers of the AMD Display driver have specifically highlighted code 
duplication as a significant impediment to their work.

@@

ArKanjo, our proposed tool, is a Command Line Interface (CLI) application designed to
help developers identify code clone duplication at the function level. Released under the
MIT license, ArKanjo is available at the url: \url{https://github.com/LipArcanjo/arkanjo}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{diagrama_mestrado}
\caption{Architecture diagram demonstrating the relationship between the tool components}
\label{fig:diagram}
\end{figure}

The tool operates in two main parts: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} is responsible for performing heavy computations to find the code 
duplications across a codebase and produce artifacts to consult information related to the 
duplications in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
by the Preprocessor to execute the tool functionalities as requested. This choice enables the 
tool to concentrate on the heavy and slow parts that can be executed only once per codebase, 
thus allowing a fast performance for multiple queries related to a single codebase.

The Preprocessor contains two main components, the Function Breaker and the Code Duplication Finder. 
The flow of how the Preprocessor works is as follow: Function Parser receives the codebase the user is interested in,
extracts the functions of the codebase along with metadata, and creates a new temporary codebase where the functions extracted become new code files. 
The Code Duplication Finder iterates over every pair of files in the Temporary Codebase, checks if they are code clones, and, 
if so, saves them in the Code Duplication Database, which is a text file that stores every code duplication as a triple 
\textbf{<function1,function2,similarity>}, where \textbf{function1} and \textbf{function2} are the functions that are duplicates 
of each other, and \textbf{similarity} is the metric given by the code duplication detection method utilized in the Code Duplication Finder.
Figure \ref{fig:diagram} illustrates the tool architecture.

The Query Responder consumes the Temporary Codebase and the Code Duplication Database to extract duplicated 
functions-related information per user request. If the user executes the Query Responder without executing the 
Preprocessor, the Query Responder calls the Preprocessor to create the required artifacts.

The Function Parser receives the input codebase and transforms it into the temporary
codebase. The Function Parser iterates through every source code file from the programming 
language it supports and uses a specific programming language function extractor
to extract every function in the file. For each function extracted, two new source code files
and a metadata file are created in the temporary codebase, represented by the pair <file
name, function name>, the source code file of the function, and the proper function.
The first new source code file contains the function’s body from the function it represents,
while the other new source code file contains the function’s signature. The metadata file
contains additional relevant information about the function, such as the function name,
the line where the function signature starts in the source code file, and the line where the
function’s body ends in the source code file. The programming languages supported at
the moment are C and Java, although Java has limited support.

The Code Duplication Finder iterates through every pair of source code files in the
temporary codebase, representing functions from the input codebase. For each pair of
files, we execute a code duplication detection method that computes a metric measuring
how similar the pair of files is, which we refer to as similarity. If the similarity is greater
than or equal to the minimum similarity threshold (a parameter provided by the user
during the Preprocessor), we store this pair of functions along with its similarity in the
Code Duplication Database. 
We implemented two methods for code duplication detection in the tool, which the user
can choose to use. The first method is based on text similarity, and the second is simpler
and based on the number of equal lines. The experiments and tests in this research were
done using the text similarity method.

@@

For the text similarity method, we treat the source code files as text and apply the
TF-IDF vector embedding method implemented by the Gensim library \cite{gensim},
then compute cosine similarity as the similarity metric. This method was chosen for 
its claimed performance, programming language independence, and the fact that it does 
not require compilable code, which is expected in the
temporary codebase, as it does not contain complete code artifacts.

For the difference method, 
we implemented a method that considered the number of exactly equal lines between two functions.
For two functions \textit{function1} and \textit{function2}, we compute the similarity as the 
ratio of duplicated lines between the two functions by the total number of lines in both functions.
This code duplication method is considerably simpler and more explainable than the text similarity method, 
as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
the \textit{diff} command built-in in the Linux environment.

@@

We propose three main functionalities for the user in the tool, called the Duplication
Explorer, Function Information, and Duplication Report. Each functionality accepts
specific parameters to perform its operations.

The Duplication Explorer is the primary functionality of our tool, designed to present
the user with pairs of duplicated functions identified by the tool. We implement optional
filters to facilitate more complex queries. Figure \ref{fig:explorer_ex} demonstrates an
example of this functionality in use.

@@

The Function Information functionality provides detailed information about a specific function. 
The functionality receives a target function from the user
and returns information such as the relative path, function name, and line numbers where
the function is defined. Additionally, it provides similar information for every function
that duplicates the given function. Figure \ref{fig:function_ex} demonstrates an
example of this functionality in use.

The Duplication Report functionality provides an overview of
duplicated code within the input codebase. This functionality calculates the number of
duplicated lines per folder in the codebase and presents the information to the user in
a readable format. Figure \ref{fig:relatory_ex} demonstrates an example of this functionality in use.

@@

>>>>

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Luan Ícaro Pinto Arcanjo}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil\\
Email: luanicaro@usp.br}
\and
\IEEEauthorblockN{David Tadokoro}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil\\
Email: davidbtadokoro@usp.br}
\and
\IEEEauthorblockN{Paulo Meirelles}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil\\
Email: paulormm@ime.usp.br}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Linux kernel massive scale (+28 M LoC, +20 K contributors) presents unique 
maintenance challenges. Surprisingly, code duplication remains a persistent issue 
in the kernel codebase, which can hinder its evolution and patching. Academic 
approaches often focus on pairwise comparison of code artifacts, which are not directly applied 
to comprehensive codebase analyses. Other existing free software tools explored in 
practice frequently suffer from limited functionality, such as primitive textual 
matching, prove too narrow in scope, or fail to deliver effective results on complex, 
large-scale codebases. Existing solutions generally fail to address the Linux kernel 
specific needs: (1) scalability to handle its size, (2) actionable results for 
developers, and (3) integration with kernel development workflows.
%
This paper presents \textit{ArKanjo}, a novel command-line tool for Linux kernel maintenance 
designed to detect and analyze function-level duplications. Released under the MIT 
license, ArKanjo employs a two-stage architecture consisting of a Preprocessor and 
a Query Responder that separates computationally intensive analysis from efficient 
querying for duplications within large codebases. Pivotal advantages of ArKanjo over 
existing solutions include: (1) optimization for C codebases with kernel-specific 
patterns; (2) preprocessing that enables rapid queries without redundant analysis; 
and (3) prioritization of duplicates that impact maintainability, such as copied 
buggy logic. 
%
We evaluate ArKanjo against real-world duplication cases in recent 
kernel versions, demonstrating its effectiveness in identifying problematic clones 
that generic tools often overlook. By identifying well-defined, manageable 
duplication instances, ArKanjo effectively lowers the barrier for new contributors, 
a capability evidenced by its role in guiding students to make their first code 
improvements to the kernel. ArKanjo offers immediate value to kernel maintainers 
and serves as a replicable model for clone detection in other large-scale free 
software projects.

@@

The Linux kernel is a foundational free software project 
critical to a significant portion of the world digital infrastructure. Maintaining 
the kernel is an enormous undertaking involving over 28 million lines of code 
and contributions from over twenty thousand developers. Within this context, code 
duplication persists as a significant challenge, a known harmful practice that 
negatively affects code readability and increases the likelihood of introducing
bugs~\cite{harmone,harmtwo}. This issue is particularly acute in the 
kernel device drivers, which comprise over 66\% of the source code~\cite{marcelo}. 
For instance, maintainers of the AMD Display driver have specifically highlighted code 
duplication as a significant impediment to their work.

@@

ArKanjo, our proposed tool, is a Command Line Interface (CLI) application designed to
help developers identify code clone duplication at the function level. Released under the
MIT license, ArKanjo is available at \textit{\url{github.com/LipArcanjo/arkanjo}}.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{diagrama_mestrado}
\caption{Architecture diagram demonstrating the relationship between the tool components}
\label{fig:diagram}
\end{figure}

The tool operates in two main parts: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} performs heavy computations to find code 
duplications across a codebase and produces artifacts to store duplication-related information 
in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
by the Preprocessor to execute the tool functionalities as requested. This design allows the 
tool to perform the heavy and time-consuming steps only once per codebase, enabling fast performance 
for multiple queries on the same codebase.

The Preprocessor contains two main components: the Function Breaker and the Code Duplication Finder. 
The workflow of the Preprocessor is as follows: the Function Parser receives the input codebase,
extracts the functions along with metadata, and creates a temporary codebase where each extracted 
function becomes a new code file. The Code Duplication Finder then iterates over every pair of files 
in the temporary codebase, checks whether they are code clones, and, if so, stores the result in 
the Code Duplication Database. This database is a text file that records each code duplication as a triple 
\textbf{<function1, function2, similarity>}, where \textbf{function1} and \textbf{function2} are the 
duplicated functions, and \textbf{similarity} is the score returned by the duplication detection method 
used in the Code Duplication Finder. Figure~\ref{fig:diagram} illustrates the tool architecture.

The Query Responder uses the temporary codebase and the Code Duplication Database to extract 
information about duplicated functions based on the user's request. If the user executes the 
Query Responder without previously executing the Preprocessor, the Query Responder automatically 
calls the Preprocessor to generate the required artifacts.

The Function Parser receives the input codebase and transforms it into the temporary
codebase. It iterates through every source code file written in a supported programming 
language and uses a language-specific extractor to isolate each function. For each extracted 
function, two new source code files and a metadata file are created in the temporary codebase, 
represented by the pair \textbf{<file name, function name>}, the source code of the function, 
and the function metadata. One of the new files contains the function body, and the other 
contains the function signature. The metadata file includes additional relevant information,
such as the function name, the line where the signature starts, and the line where the function 
body ends. Currently, the supported programming languages are C and Java, although Java support is limited.

The Code Duplication Finder iterates through every pair of source code files in the
temporary codebase, each representing a function from the input codebase. For each pair,
a code duplication detection method computes a similarity score. If the similarity is greater 
than or equal to the minimum threshold (provided by the user during preprocessing), the pair 
is stored in the Code Duplication Database along with the similarity score.

We implemented two duplication detection methods in the tool, which users can choose 
between. The first method is based on text similarity, and the second is a more straightforward approach 
based on the number of identical lines. The experiments and tests in this research were 
conducted using the text similarity method.

@@

For the text similarity method, we treat the source code files as plain text and apply the
TF-IDF vector embedding method implemented by the Gensim library~\cite{gensim},
then compute cosine similarity as the similarity metric. We chose this method for 
its reported performance, programming language independence, and the fact that it does 
not require compilable code, an important aspect given that the
temporary codebase does not contain complete code artifacts.

We implemented a more straightforward approach for the difference method,
considering the number of exactly equal lines between two functions.
For two functions, \textit{function1} and \textit{function2}, we compute the similarity as the 
ratio of duplicated lines to the total number of lines across both functions.
This method is considerably simpler and more explainable than the text similarity method, 
The metric is defined by the proportion of common lines between the two functions. 
We use the \textit{diff} command built into the Linux environment to compute the number of equal lines.

@@

We propose three main functionalities for the user in the tool, called Duplication
Explorer, Function Information, and Duplication Report. Each functionality accepts
specific parameters to perform its operations.

The Duplication Explorer is the primary functionality of our tool, designed to present
the user with pairs of duplicated functions identified by the tool. We implement optional
filters to support more complex queries. Figure~\ref{fig:explorer_ex} demonstrates an
example of this functionality in use.

@@

The Function Information functionality provides detailed information about a specific function. 
It receives a target function from the user
and returns information such as the relative path, function name, and line numbers where
the function is defined. Additionally, it provides similar information for every function
that duplicates the given function. Figure~\ref{fig:function_ex} demonstrates an
example of this functionality in use.

The Duplication Report functionality provides an overview of
duplicated code within the input codebase. It calculates the number of
duplicated lines per folder in the codebase and presents the information to the user in
a readable format. Figure~\ref{fig:relatory_ex} demonstrates an example of this functionality in use.

@@

\section*{Acknowledgment}

This study was financed, in part, by  the University of
São Paulo/USP (Proc. 22.1.9345.1.2), the São Paulo Research Foundation/FAPESP
and the São Paulo State Data Analysis System Foundation/SEADE (Proc. 2023/18026-8), Brazil.
